---
title: "M1-3 DS"
author: "Mei Gilhousen"
date: "9/14/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

plots
Read in Dataset and filter for target state and separate by population
There are 25 observations after filtering with the 7 original columns all remaining
```{r}
df<-read.csv("C:\\Users\\Student\\Downloads\\df.csv")
va_stem<-subset(df, df$state_name=="Virginia" )
as.data.frame(va_stem)
va_stem

```
Graphing data using ggplot to create a scatterplot comparing the STEM Population and early career pay, colored by percent of STEM students (shown on the legend). The graph shows a positive, linear, and moderately strong correlation between the two variables shown by the trend line. The confidence level is higher where the population is between 0% and 20%, since there are more data points compared to the low number of observations from the 20% to 40% population. 
```{r}
library(ggplot2)

ggplot(va_stem ,               
       aes(x = stem_percent,           
           y = early_career_pay,
           color=stem_percent
          )) +       
  geom_point() +
  geom_smooth(color="pink", fill="white") + 
  labs(
    title = 'Salary potential and STEM distribution',
    subtitle = 
    ) +
  ylab('early career pay in dollars') +
  xlab('stem distribution in %') +
  labs()
  

```

The next plot measures the same two variables of early career pay and STEM populations at each school in VA. The boxplot below compares the two groups of STEM populations greater than or less than 20%. Key points are that the median salary for the less than 20% group is about $50000, and the greater than 20% group has a median salary of about $58000. This is evidence that helps support our hypothesis. Looking at the two groups, we can also see that the spread for the low STEM population group is higher. There are no outliers shown by the graphs. 
```{r}
ggplot(va_stem, aes(stem_percent>20, early_career_pay, fill=stem_percent)) +
  geom_boxplot()+
  labs(
    title = 'Salary Potential and STEM Distribution in VA Colleges',
    subtitle = "Boxplot"
    ) +
  ylab('Early Career Pay in Dollars') +
  xlab('STEM Population >20%') +
  labs()
  
  
  
```
Descriptive statistics for Average and SD of Early Career Pay.
Next, we computed the average pay and standard deviation for pay for each group. Although the averages support our hypothesis, the standard deviations are somewhat high.
```{r}
library(dplyr)

va_stem %>%
  select(stem_percent, early_career_pay) %>%
  group_by(stem_percent>20) %>%
  summarise(avg = mean(early_career_pay),
            sd = sd(early_career_pay)) -> salary_table
salary_table
            
```

Calculating R-Squared Value of the correlation and significance. 
Finding the correlation coefficient comes out to 59%. 59% of the variation in the early career pay variable is explained by the STEM population at each school. 

Testing Significance
1. Null hypothesis: variable 1 and variable have a correlation equal to 0.
2. Alternative hypothesis: variable 1 and variable 2 have a correlation not equal to 0.

Running a significance test on the two variables, we found that the p-value was .001983. At a 95% confidence level, we can reject the null hypothesis that the correlation is equal to 0 as the p-value is statistically significant. The confidence interval is above 0, which supports the decision to reject the null hypothesis as well, although the range is quite large, likely due to the limited number of observations in the dataset.

In conclusion, there is a significant relationship between early career pay and STEM populations in Virginia colleges and universities. 
```{r}
correlation_coef<-cor(va_stem$stem_percent, va_stem$early_career_pay)
correlation_coef

cor.test(va_stem$stem_percent, va_stem$early_career_pay)
cor.test
```







